---
title: "Week 1 Problem Set"
subtitle: "Avogadro's Law"
author: "Stephanie Plaza-Torres"
date: "Last knitted `r format(Sys.Date(), '%d %b %Y')`" 
output:
  html_document: 
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: true
    toc_depth: 3
    code_folding: show
editor_options:
  chunk_output_type: inline
---
<p>&nbsp;</p>
# ***Avogadro's Law***
<p>&nbsp;</p>

## Learning goals

After doing this exercise you should be able to:

1. Define what each variable of Avogadro's Law stands for.
2. Derive the $k$ constant from the Ideal Gas Law.
3. Show that the volume of a gas is directly proportional to the number of moles of gas present in the sample.
4. Calculate the volume of gas given a certain quantity of moles (when pressure and temperature are constant).

## Background

*Avogadro's Law was hypothesized by Amedeo Avogadro in 1812.*

Avogadro's Law states that:

> Equal volumes of any gas, at the same temperature and pressure, contain the same number of molecules.

In this statement the **number of molecules** refers to the **number of moles**. Moles measure the amount of a substance. A mole's particles can be be represented by things like the atoms of a substance or the molecules of a substance. Although this is not very important for our lesson today, each mole of substance contains $6.02214076 × 10^{23}$ particles.

>This implies that the volume of a gas, kept at a constant temperature and pressure, is directly proportional to the number of moles of gas present in the sample. 

*After this point in the lesson we'll use this notation: $n$ is the number of moles of gas and $V$ is the volume of gas in liters.*

You can express Avogadro's Law mathematically.

- The volume of a gas is directly proportional to the moles of that gas:

$$
V\ \propto n\
$$
- Which can also be expressed like:

$$
\begin{align}
\frac{V}{n} & = k\\
\\
V & = k \times n\
\end{align}
$$
or 

$$
\frac{V_{1}}{n_{1}}\ = \frac{V_{2}}{n_{2}}\ 
$$
where $k$ is a constant that can be calculated.

<p>&nbsp;</p>

## Examples

You can calculate the $k$ constant by deriving it from the ideal gas law equation where $P$ is pressure in pascals, $R$ is the gas constant, and $T$ is the Kelvin temperature:

$$
PV\ = nRT\
$$
- If you solve for $\frac{V}{n}$ you'll have:

$$
\begin{align}
\frac{V}{n}\ = \frac{RT}{P}\\
\\
k = \frac{RT}{P}\\
\end{align}
$$
which shows that the $k$ constant is for a fixed temperature and pressure of a gas.

<p>&nbsp;</p>

So, given the formula above let's calculate $k$:

```{r include = FALSE}
calc_k_list <- list(R=8.314, T=80, P=20)
names(calc_k_list)
```

If $R$ is 8.314 J/mol⋅K, $T$ is 80 Kelvin, and $P$ is 20 Pascals:
$$
k = \frac{8.314 \times 80}{20}\\
$$

```{r}
k <- (calc_k_list$R*calc_k_list$T)/calc_k_list$P
```

$k$ equals 
```{r}
print (k)
print ("J/mol⋅Pascal")
```
<p>&nbsp;</p>

If we use the above $k$ constant we can show that if the moles of gas increase, in the constant $T$ and $P$ shown above, the volume increases. Let's apply the $k$ from above to calculate what would be the volume of 1 mole of oxygen gas versus 2 moles of oxygen gas. If we use the equation $V = k \times n$: 

If we have 1 mol of oxygen gas the volume would be:

$$
\begin {align}
V & = 33.256 \times 1\ \\
V & = 33.256\ liters
\end{align}
$$
But if we have 2 mol of oxygen gas the volume would be:

$$
\begin {align}
V & = 33.256 \times 2\ \\
V & = 66.512\ liters
\end{align}
$$
______________________________________________________________________________________

### Another example

If 3 mol of hydrogen gas occupies a volume of 6 liters, then how many liters would 6 mol of hydrogen gas occupy?

We know $V_{1}$ is 6 liters, that $n_{1}$ is 3 mol, and that $n_{2}$ is 6 mol. We are missing our final volume which is $V_{2}$. If we solve for $V_{2}$:

$$
\begin{align}
\frac{6\ liters}{3\ mol}\  = \frac{V_{2}}{6\ mol}\\
\\
V_2 = \frac{6\ liters\times 6\ mol}{3\ mol}\\
\\
\end{align}
$$


```{r}
#if V1 = 6, n1 = 3, n2 = 6, and we are missing V2, and we'll solve for V2 using the equation formatted above. First let's set up our variables:

V1 <- 6
n1 <- 3 
n2 <- 6

cat ("V1 is ", V1,"\n")
cat ("n1 is ", n1,"\n")
cat ("n2 is ", n2,"\n")

```

```{r}
#Once we set up our variables let's solve the equation using R operators and the variables we established:

result <- (V1*n2)/n1
print(result)
```
So the value we obtained was that:
$$
V_2 = 12\ liters
$$
We can see that $V_{2}$ is larger than $V_{1}$ because it contains more moles of gas.

<p>&nbsp;</p>

## Questions

Q1. What does $V_{2}$ mean in the context of this equation:
$$
\frac{V_{1}}{n_{1}}\ = \frac{V_{2}}{n_{2}}\ 
$$

Q2. If you have a 22 liters of oxygen gas that contain 3 mol of oxygen, how many mol of gas would 30 liters of oxygen have? These gases would all be at a constant STP (which is the Standard Temperature and Pressure, $273.15$ $K$ and $10^5$ $Pa$)

Q3. What happens (to the mol value) if the volume of a gas increases and the temperature and pressure remain constant?

Q4. What are the two variables that should remain constant to comply with Avogadro's Law?
