---
title: "Week 2 Problem Set"
subtitle: "Ohm's Law"
author: "Stephanie Plaza-Torres"
date: "Last knitted `r format(Sys.Date(), '%d %b %Y')`" 
output:
  html_document: 
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: true
    toc_depth: 3
    code_folding: show
editor_options:
  chunk_output_type: inline
---
<p>&nbsp;</p>
# ***Ohm's Law***
<p>&nbsp;</p>

## Learning goals

After doing this lesson you should be able to:

1. Define what each variable of the Ohm's Law equation stands for.
2. Create a tibble dataframe, use the dataframe for creating plots with ggplot2, and use the dataframe to perform calculations for variables in the Ohm's Law equation.
3. Visualize Ohm's Law graphically.

## Background

Ohm's Law is a concept that focuses on the study of electrical circuits. The equation for Ohm's Law is:

$$
R = \frac{V}{I} 
$$
In this equation:

> $I$ is current and it is measured in Amperes.
> $V$ is the electric potential difference between two points and it is measured in Volts.
> $R$ is a measurement of how a material oposes the flow of an electric current and it is measured in Omhs. 
> You can re-express the equation by rearranging the variables above.

$R$ is controlled by all of the resistors between the two measured points used to get the electric potential difference. 

<p>&nbsp;</p>

## Examples

### Graphical Visualization with ggplot2 and creating tibble dataframes

You can represent the Ohm's Law equation *graphically*:

```{r Ohm Law Graph, message=FALSE}
# Creating my dataframe

# For this dataframe we will make a table with six voltage values and six current values. The dataframe will have a V column and an I column, and six rows.


library(dplyr) # Load the dplyr library to use the tibble function.
V_vs_I <- # Assigns what is calculated by the tibble function to our dataframe, V_vs_I.
  tibble( # Calls the tibble function
    y_axis_V = seq(0.5, 3, by = 0.5), # creates y_axis_V column that contains V values. seq creates a sequence of six numbers that starts with 0.5 and ends with 3, with an increment of 0.5 from each number. So, 0.5, 1.0, 1.5...
    x_axis_I = seq(0.005, 0.030, by = 0.005) # creates x_axis_I column that contains I values. seq creates a sequence of six numbers that starts with 0.005 and ends with 0.030, with an increment of 0.005 from each number. So, 0.005, 0.010, 0.015...
  ) 


# Checking my dataframe is correct

V_vs_I # Prints V_vs_I dataframe


# Graphing Voltage vs. Current

library(ggplot2) # Loads ggplot2 library which allows us to use the ggplot function, which creates graphs. 
line_graph <- V_vs_I %>% # Stores our graph into line_graph and forwards V_vs_I values to ggplot function
  ggplot() + aes(x = x_axis_I, y = y_axis_V) + geom_line() + labs(title="Voltage vs. Current",
        x ="Current (Amperes)", y = "Voltage (Volts)")# Calls ggplot function, establishes values to be used for x axis and y axis, establishes that you want a line plot, and defines the graph's labels. If you wrote geom_point() you would get a scatter plot.

print (line_graph)
```
<p>&nbsp;</p>

### Calculating Resistance using our existing tibble dataframe and showing R is constant

If we use the data from the tibble dataframe we created above we can calculate $R$ for every correspondent $V$ and $I$ value. You will notice below, that the $R$ remains constant which matches the slope represented by the graph we created above. If we get the slope of the line graph above we will get the same $R$ value that we get by using the Ohm's Law equation:
$$
R = \frac{V}{I} 
$$
If point 2 of the line graph we plotted are the $V$ and $I$ values of the sixth row of our dataframe and point 1 of the line graph we plotted are the values from the first row of our data frame. We can calculate the slope of the line using this equation:
$$
R = \frac{V_{2}-V_{1}}{I_{2}-I_{1}}\
$$
which would be:

$$
R = \frac{3-0.5}{0.030-0.005}\ = 100 \hspace{0.1cm} Ohms
$$
And if we use the Ohm's Law equation to calculate $R$ using our tibble dataframe we can see that it is the same as the one calculated above (using the slope):
```{r Showing R is Constant, message=FALSE}

V_vs_I$R <- V_vs_I$y_axis_V/V_vs_I$x_axis_I # V_vs_I$R creates a new column titled R. <- assigns what is calculated on the right besides it to V_vs_I$R. V_vs_I$y_axis_V/V_vs_I$x_axis_I calls the y_axis_V column values of our dataframe and divides them by the x_axis_I values of our dataframe.

```

This calculation emulates this equation: 
$$
R = \frac{V}{I} 
$$

```{r, message=FALSE}
V_vs_I # Prints the V_vs_I dataframe where R shows the calculated resistance
```

<p>&nbsp;</p>
___________________________________________________________________________________________________________

## Questions

1. In Ohm's Law $R$ represents the _____________ (fill in the blank) in Ohms.

2. Rearrange the Ohm's Law equation to solve for $I$. What would this equation imply? Explain the variables in the equation. 

3. What is the kind of relationship between $V$ and $I$ graphically for Ohm's Law? What does the slope represent?

4. Apply your dataframe and ggplot2 knowledge. Create a dataframe using this rearranged version of Omh's Law $V=IR$. The dataframe's $R$ variable should start at 1 and end at 6, incrementing by 1 from one number to another. The $I$ variable should go from 0 to 1000 incrementing by 200. Calculate $V$ for every row in your dataframe using the equation above. Add the results of $V$ as a new column in your dataframe. Use your new $V$ values alongside the $I$ values to make a plot showing the relationship $R=V/I$.


