---
title: "Week 2 Answer Key"
subtitle: "Ohm's Law"
author: "Stephanie Plaza-Torres"
date: "Last knitted `r format(Sys.Date(), '%d %b %Y')`"
output: html_document

---
# _Questions and Answers_

## Questions

1. In Ohm's Law $R$ represents the _____________ (fill in the blank) in Ohms.

2. Rearrange the Ohm's Law equation to solve for $I$. What would this equation imply? Explain the variables in the equation. 

3. What is the kind of relationship between $V$ and $I$ graphically for Ohm's Law? What does the slope represent?

4. Apply your dataframe and ggplot2 knowledge. Create a dataframe using this rearranged version of Omh's Law $V=IR$. The dataframe's $R$ variable should start at 1 and end at 6, incrementing by 1 from one number to another. The $I$ variable should go from 0 to 1000 incrementing by 200. Calculate $V$ for every row in your dataframe using the equation above. Add the results of $V$ as a new column in your dataframe. Use your new $V$ values alongside the $I$ values to make a plot showing the relationship $R=V/I$.



### Question 1

1. In Ohm's Law $R$ represents the _____________ (fill in the blank) in Ohms.

### Answer 1

Resistance

______________________________________________________________________________________

### Question 2

2. Rearrange the Ohm's Law equation to solve for $I$. What would this equation imply? Explain the variables in the equation. 

### Answer 2

$$
I = \frac{V}{R} 
$$
It would imply that the current of a system is represented by change in voltage between two points of the system divided by the resistance of the whole system.

> $I$ is current and it is measured in Amperes.
> $V$ is the electric potential difference between two points and it is measured in Volts.
> $R$ is a measurement of how a material oposes the flow of an electric current and it is measured in Omhs.
______________________________________________________________________________________

### Question 3

3. What is the kind of relationship between $V$ and $I$ graphically for Ohm's Law? What does the slope represent?

### Answer 3

The relationship between $V$ and $I$ graphically is a linear relationship. The slope of this line represent the resistance of the system.

______________________________________________________________________________________

### Question 4

4. Apply your dataframe and ggplot2 knowledge. Create a dataframe using this rearranged version of Omh's Law $V=IR$. The dataframe's $R$ variable should have 6 values that are all 10 (use the rep function, see help("rep") to understand how to use it). The $I$ variable should go from 0 to 1000 incrementing by 200. Calculate $V$ for every row in your dataframe using the equation above. Add the results of $V$ as a new column in your dataframe. Use your new $V$ values alongside the $I$ values to make a plot showing the relationship $R=V/I$.


### Answer 4

```{r Q4 Answer Dataframe, message=FALSE}
# Creating my dataframe

# For this dataframe we will make a table with six voltage values and six current values. The dataframe will have a V column and an I column, and six rows.


library(dplyr) # Load the dplyr library to use the tibble function.
dataframe <- # Assigns what is calculated by the tibble function to our dataframe.
  tibble( # Calls the tibble function
    R = rep(10, each = 6),
    I = seq(0, 1000, by = 200)
  ) 


```

```{r Q4 Answer Plot, message=FALSE}
# Calculating V

dataframe$V <- dataframe$I*dataframe$R

# Checking my dataframe is correct

dataframe # Prints dataframe

# R = V/I graph

library(ggplot2) # Loads ggplot2 library which allows us to use the ggplot function, which creates graphs. 
VI_graph <- dataframe %>% 
  ggplot() + aes(x = I, y = V) + geom_line() + labs(title="Voltage vs. Current",
        x ="Current (Amperes)", y = "Voltage (Volts)")# Calls ggplot function, establishes values to be used for x axis and y axis, establishes that you want a line plot, and defines the graph's labels. If you wrote geom_point() you would get a scatter plot.

print (VI_graph)

```