---
title: "Week 4 Answer Key"
subtitle: "Introduction to Phylogenetic Trees"
author: "Stephanie Plaza-Torres"
date: "Last knitted `r format(Sys.Date(), '%d %b %Y')`"
output: html_document

---
# _Questions and Answers_


## Question 1

1. Look at the `sample_tree_data.xlsx` file. Edit the file to create your own presence/absence matrix (it should have 6 traits and 6 organisms), and name it something with your initials at the end. Once you do this, load the file into R, make it a dataframe, asssign it to a variable, and put column and row headers to it. The dataframe should be named `my_matrix`. 


### Answer 1

```{r message = FALSE}
library(tidyverse)
library(readxl)

mymatrix <- data.frame(read_excel("data/sample_tree_data_SPT.xlsx"))
colnames(mymatrix) <- c("Trait 1", "Trait 2", "Trait 3", "Trait 4", "Trait 5", "Trait 6") 
rownames(mymatrix) <- c("A", "B", "C", "D", "E", "F") 

mymatrix
```

______________________________________________________________________________________

## Question 2

2. What do phylogenetic trees illustrate? What is a clade?

### Answer 2

Phylogenetic trees show the evolutionary relationships between a set of organisms or taxa. Every end tip of a tree represents a taxa, and the nodes that join the branches show a common ancestor to those taxa. A clade is a part of a phylogenetic tree that shows a group of organisms that includes an ancestor of that organism and all of its descendants.
______________________________________________________________________________________

## Question 3

3. Construct a presence/absence dataframe (you don't need to code it into R) that would create the first example phylogenetic tree of the background section of this lesson.

### Answer 3

The matrix below would be the simplest possible matrix for this question. An alternate matrix could include up to 4 additional traits if the student decides to assign a new trait (one per branch) that characterizes branches A, B, C, or D. Each of these orgasnisms would have their own particular trait in this case.

```{r message = FALSE}
trait1 <- c(1,1,1,1)
trait2 <- c(1,1,0,0)
trait3 <- c(0,0,1,1)

mydf <- data.frame(trait1, trait2, trait3)

colnames(mydf) <- c("Trait 1", "Trait 2", "Trait 3") 
rownames(mydf) <- c("A", "B", "C", "D") 

mydf
```
 
## Question 4

4. Change the contents of the data in the third column and third row of your `mymatrix` dataframe to either "YES" or "NO", where "YES" would be for a 1 value and "NO" would be for a 0 value. 

### Answer 3

```{r}
mymatrix[3,3] <- "YES"
mymatrix[3,3] <- "NO"
```

